---
title: "Transit in LA and Chicago"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tidycensus)
library(here)
library(sf)
library(ggpubr)
library(plotly)
```

```{r global, include=FALSE}
# create data directory if it doesn't exist
if(!dir.exists(here("data"))){
  dir.create(here("data"))
}

# PUMA data
if(!file.exists(here("data", "CA_2019_transit.rds"))){
  get_acs_transit <- function(year_input, state_input){
    # use tidycensus API to extract transit users and total worker pop
    get_acs(geography = "public use microdata area",
            variables = paste("B08006", sprintf("%03d", c(1,2,8,17)), sep = "_"),
            year = year_input,
            state = state_input,
            geometry = TRUE,
            survey = "acs1") %>% 
      # change variables names to be more descriptive
      mutate(
        variable = str_replace_all(variable, "B08006_001", "total_workers"),
        variable = str_replace_all(variable, "B08006_002", "car_users"),
        variable = str_replace_all(variable, "B08006_008", "transit_users"),
        variable = str_replace_all(variable, "B08006_017", "work_from_home")) %>%
      # save as RDS file
      saveRDS(file = here("data", paste0(paste(state_input, year_input, "transit", sep = "_"), ".rds")))
  }

  # iterate over combinations of states/years
  state_years <- expand_grid(states = c("CA", "IL"),
                             years = c(2019, 2021))
  walk2(.x = state_years$years, .y = state_years$states, get_acs_transit)
}

# load datasets
CA_2019 <- readRDS(here("data", "CA_2019_transit.rds"))
CA_2021 <- readRDS(here("data", "CA_2021_transit.rds"))
IL_2019 <- readRDS(here("data", "IL_2019_transit.rds"))
IL_2021 <- readRDS(here("data", "IL_2021_transit.rds"))

# -------------------------------------------------------
# -------------------------------------------------------

# county-wide data
if(!file.exists(here("data", "county_transit.rds"))){
  get_county_transit <- function(year_input, state_input, county_input){
    get_acs(geography = "county",
            variables = paste("B08006", sprintf("%03d", c(1,2,8,17)), sep = "_"),
            year = year_input,
            state = state_input,
            county = county_input,
            survey = "acs1") %>% 
      # change variables names to be more descriptive
      mutate(
        variable = str_replace_all(variable, "B08006_001", "total_workers"),
        variable = str_replace_all(variable, "B08006_002", "car_users"),
        variable = str_replace_all(variable, "B08006_008", "transit_users"),
        variable = str_replace_all(variable, "B08006_017", "work_from_home"),
        year = year_input
      )
  }
  county_years <- expand_grid(states = c("CA", "IL"),
                             years = c(2019, 2021)) %>% 
                 cbind(counties = c("Los Angeles", "Los Angeles", "Cook", "Cook"))
  pmap(.l = list(county_years$years, 
                 county_years$states, 
                 county_years$counties), 
       .f = get_county_transit) %>% 
  bind_rows() %>%
  saveRDS(file = here("data", "county_transit.rds"))
}

county_transit <- readRDS(here("data", "county_transit.rds"))

#--------------------
#--------------------

# calculate raw change/change in percent
calculate_change <- function(list_data, filtering_string){
  list_data %>% 
    # filter out county from state PUMAs
    map(
      ~filter(.x, str_detect(NAME, filtering_string)) %>%
        select(-moe)
    ) %>%
    # add year info to variable names in prep for pivot
    map2(.y = c("2019", "2021"),
         .f = ~mutate(.x, variable = str_c(variable, "_", .y))
    ) %>% 
    bind_rows() %>% 
    pivot_wider(
      names_from = "variable",
      values_from = "estimate"
    ) %>% 
    # raw_change: absolute change in ridership
    # perc_change: percent transit use in 2021 - percent transit use in 2019
    mutate(
      # transit
      raw_change_transit = transit_users_2021 - transit_users_2019,
      perc_transit_2019 = transit_users_2019/total_workers_2019 * 100,
      perc_transit_2021 = transit_users_2021/total_workers_2021 * 100,
      perc_change_transit = perc_transit_2021 - perc_transit_2019,
      # car
      raw_change_car = car_users_2021 - car_users_2019,
      perc_car_2019 = car_users_2019/total_workers_2019 * 100,
      perc_car_2021 = car_users_2021/total_workers_2021 * 100,
      perc_change_car = perc_car_2021 - perc_car_2019,
      # car
      raw_change_wfh = work_from_home_2021 - work_from_home_2019,
      perc_wfh_2019 = work_from_home_2019/total_workers_2019 * 100,
      perc_wfh_2021 = work_from_home_2021/total_workers_2021 * 100,
      perc_change_wfh = perc_wfh_2021 - perc_wfh_2019,
    )
}

chicago_transit <- calculate_change(list(IL_2019, IL_2021), "(Chicago City)|(Cook County)")
la_transit <- calculate_change(list(CA_2019, CA_2021), "Los Angeles County")

rm(CA_2019, CA_2021, IL_2019, IL_2021)
```


About {data-orientation=rows}
==================

**What is the purpose of this dashboard?**

The purpose of this dashboard is to compare changes in public transit use (and other forms of transportation) over the pandemic in Los Angeles, CA and Chicago, IL.

**Where do the data come from?**

The datasets used in this dashboard are from the [1-Year American Community Survey (ACS)](https://www.census.gov/data/developers/data-sets/acs-1year.html) in [2019](https://www.census.gov/newsroom/press-kits/2020/acs-1year.html) and [2021](https://www.census.gov/newsroom/press-kits/2022/acs-1year.html). The data were obtained from the ACS using the [tidycensus](https://walker-data.com/tidycensus/) API in R.

Row
------------------

### Navigating the dashboard

**Here's a two-minute overview of this dashboard:**\n

<iframe width="560" height="315" src="https://www.youtube.com/embed/uBZ712Bv5jo?si=IE5FuScKlgTud7FL" frameborder="0" data-external="1"   allowfullscreen></iframe>


The Data
==================

Column {.tabset .tabset-fade}
------------------

### Los Angeles 2019/2021 (PUMA-level)

This dataset contains estimates of various transportation mode uses (car, public transit, and work from home) by 2010 Census [Public Use Microdata Area (PUMA)](https://www.census.gov/programs-surveys/geography/guidance/geo-areas/pumas.html#overview) in Los Angeles County, CA in 2019 and 2021.

```{r}
la_transit %>% 
  st_drop_geometry() %>% 
  DT::datatable(options = list(
    pageLength = 25
  ))
```

### Chicago 2019/2021 (PUMA-level)

This dataset contains estimates of various transportation mode uses (car, public transit, and work from home) by 2010 Census [Public Use Microdata Area (PUMA)](https://www.census.gov/programs-surveys/geography/guidance/geo-areas/pumas.html#overview) in Cook County, IL in 2019 and 2021.

```{r}
chicago_transit %>% 
  st_drop_geometry() %>% 
  DT::datatable(options = list(
    pageLength = 25
  ))
```

### County-level (2019 and 2021)

This dataset contains county-wide estimates of various transportation mode uses (car, public transit, and work from home) in Cook County, IL and Los Angeles County, CA in 2019 and 2021. 

```{r}
county_transit %>% 
  DT::datatable(options = list(
    bPaginate = FALSE
  ))
```

Countywide Commute Modality
==================

Column
------------------

### Percent Ridership by Modality

```{r add-other, include=FALSE}
# add "other category"
county_transit <- county_transit %>%
  select(-GEOID, -moe) %>% 
  split(~NAME + year) %>% 
  map(.f = ~add_row(.x, variable = "other",
                    estimate = 2*max(.x$estimate) - sum(.x$estimate))) %>% 
  map(.f = ~mutate(.x, perc_use = estimate / max(estimate)) %>% 
                fill(NAME, year)) %>% 
  bind_rows()
```

```{r}
# graph
county_transit %>% 
  group_by(NAME, year) %>% 
  mutate(
    perc_use = estimate / max(estimate),
    variable = factor(variable, 
                      levels = c("car_users", "work_from_home", "transit_users", "other"),
                      labels = c("Car", "Work from home", "Public transportation", "Other"))
  ) %>% 
  filter(variable != "total_workers") %>% 
  ggplot(aes(x = year, y = perc_use, fill = variable)) +
    geom_area() +
    facet_wrap(~NAME) +
    theme_minimal() +
    scale_x_continuous(breaks = c(2019, 2021)) +
    scale_fill_viridis_d() +
    labs(
      title = "Change in Commute Modality, 2019-2021",
      x = "Year",
      y = "% workers using __ to commute",
      fill = "",
      caption =  "Data from American Community Survey 2019 & 2021\nNote: Data is incomplete"
    ) +
  theme(legend.position = "bottom")
```

Column
------------------

### Raw Ridership by Modality

```{r}
county_transit %>% 
  group_by(NAME, year) %>% 
  mutate(
    perc_use = estimate / max(estimate),
    variable = factor(variable, 
                      levels = c("car_users", "work_from_home", "transit_users", "other"),
                      labels = c("Car", "Work from home", "Public transportation", "Other"))
  ) %>% 
  filter(variable != "total_workers") %>% 
  ggplot(aes(x = year, y = estimate, color = variable)) +
    geom_point() +
    geom_line() +
    facet_wrap(~NAME) +
    theme_minimal() +
    scale_x_continuous(breaks = c(2019, 2021)) +
    scale_y_continuous(trans = "log10", 
                       limits = c(1e5, 5e6),
                       breaks = c(1e5, 2.5e5, 5e5, 7.5e5, 1e6, 2.5e6, 5e6),
                       labels = scales::label_number(suffix = "K", scale = 1e-3)) +
    scale_color_viridis_d() +
    labs(
      title = "Change in Commute Modality, 2019-2021",
      x = "Year",
      y = "Number of workers using __ to commute",
      color = "",
      caption =  "Data from American Community Survey 2019 & 2021\nNote: Data is incomplete; log(10) y-axis used"
    ) +
  theme(legend.position = "bottom")
```

Chicago {data-navmenu="Most-Used Modalities"}
==================

```{r find-first-second, include=FALSE}
find_max <- function(data, yr){
  yr = as.character(yr)
  data %>% 
    select(NAME, contains(yr), -contains("perc"), -contains("workers")) %>% 
    pivot_longer(contains(yr), names_to = "transit_type") %>% 
    group_by(NAME) %>%
    slice_max(value, with_ties = FALSE) %>% 
    ungroup() %>% 
    mutate(transit_type = case_when(is.na(value) ~ NA, 
                                    .default = str_replace(transit_type, "_20\\d{2}", "")),
           transit_type = factor(transit_type, 
                                 levels = c("car_users", "work_from_home", "transit_users"),
                                 labels = c("Car", "Work from home", "Public transportation")
                                 ))
}

find_second <- function(data, yr){
  yr = as.character(yr)
  long_data <- data %>% 
    select(NAME, contains(yr), -contains("perc"), -contains("workers")) %>% 
    pivot_longer(contains(yr), names_to = "transit_type") %>% 
    mutate(transit_type = str_replace(transit_type, "_20\\d{2}", ""),
           transit_type = factor(transit_type, 
                                 levels = c("car_users", "work_from_home", "transit_users"),
                                 labels = c("Car", "Work from home", "Public transportation")
                                 ))
  max_data <- find_max(data, yr) %>% 
    st_drop_geometry()
  long_data %>%
    anti_join(max_data, by = join_by(NAME, transit_type)) %>% 
    group_by(NAME) %>%
    slice_max(value, with_ties = FALSE) %>% 
    ungroup() %>% 
    mutate(transit_type = case_when(is.na(value) ~ NA, 
                                    .default = transit_type),
           )
}

```

Column
------------------

### Most-used

```{r}
chi_max_2019 <- find_max(chicago_transit, 2019) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(
      subtitle = "2019",
      fill = ""
    )

chi_max_2021 <- find_max(chicago_transit, 2021) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(subtitle = "2021")

chi_max <- ggarrange(chi_max_2019, chi_max_2021, common.legend = TRUE, legend = "bottom")
annotate_figure(chi_max, 
                top = text_grob("Most-used transportation mode by PUMA\nin Cook County, 2019-2021",
                                face = "bold",
                                size = 12))
```

Column
------------------

### Second-most-used


```{r}
chi_sec_2019 <- find_second(chicago_transit, 2019) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(
      subtitle = "2019",
      fill = ""
    )

chi_sec_2021 <- find_second(chicago_transit, 2021) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(subtitle = "2021")

chi_sec <- ggarrange(chi_sec_2019, chi_sec_2021, common.legend = TRUE, legend = "bottom")
annotate_figure(chi_sec, 
                top = text_grob("Second most-used transportation mode\nby PUMA in Cook County, 2019-2021",
                                face = "bold",
                                size = 12))
```


Los Angeles {data-navmenu="Most-Used Modalities"}
==================

Column
------------------

### Most-used

```{r}
la_max_2019 <- find_max(la_transit, 2019) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(
      subtitle = "2019",
      fill = ""
    )

la_max_2021 <- find_max(la_transit, 2021) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(subtitle = "2021")

la_max <- ggarrange(la_max_2019, la_max_2021, common.legend = TRUE, legend = "bottom")
annotate_figure(la_max, 
                top = text_grob("Most-used transportation mode by PUMA\nin Los Angeles County, 2019-2021",
                                face = "bold",
                                size = 12))
```

Column
------------------

### Second-most-used


```{r}
la_sec_2019 <- find_second(la_transit, 2019) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(
      subtitle = "2019",
      fill = ""
    )

la_sec_2021 <- find_second(la_transit, 2021) %>% 
  ggplot(aes(fill = transit_type)) +
    geom_sf(color = "black") +
    coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
    scale_fill_viridis_d(na.value = "grey", drop = FALSE) +
    theme_void() +
    labs(subtitle = "2021")

la_sec <- ggarrange(la_sec_2019, la_sec_2021, common.legend = TRUE, legend = "bottom")
annotate_figure(la_sec, 
                top = text_grob("Second most-used transportation mode\nby PUMA in Los Angeles County, 2019-2021",
                                face = "bold",
                                size = 12))
```



Chicago {data-navmenu="Change in Transit Use"}
==================

Column {.sidebar}
------------------


```{r}
selectInput(
  inputId = "location_input_chi",
  label = "Choose a location:",
  choices = c("Cook County" = "(Chicago City)|(Cook County)",
              "City of Chicago" = "Chicago City",
              "Suburban Cook County" = "Cook County")
)

selectInput(
  inputId = "transit_input_chi",
  label = "Choose a transit type:",
  choices = c("Car" = "change_car",
              "Work from home" = "change_wfh",
              "Public transportation" = "change_transit")
)
```


Column
------------------

### Change in raw ridership/use

```{r}
renderPlotly({
  chicago_change <- chicago_transit %>%
    select(NAME, contains(input$transit_input_chi)) %>% 
    filter(str_detect(NAME, input$location_input_chi)) %>% 
    pivot_longer(cols = contains("change")) %>% 
    mutate(
      tooltip = NAME %>% 
        str_replace("--", "\n - ") %>% 
        str_replace(" PUMA(;|,) Illinois", "") %>% 
        str_replace_all("Townships? ?", "") %>% 
        str_replace_all(", |(?<!(North|South|East|West) )&", "\n - ")
    )
  
  p_perc <- chicago_change %>% 
    filter(str_detect(name, "raw")) %>% 
    ggplot(aes(fill = value, text = tooltip)) +
      geom_sf(color = "black", lwd = 0.1) +
      coord_sf(xlim = c(-88.3, -87.5), ylim = c(41.4, 42.2)) + 
      scale_fill_viridis_c(name = "Use", option = "magma") +
      theme_void() +
      labs(
        title = "Change in Raw Use"
      )
  
  ggplotly(p_perc, tooltip = "text") %>% 
    style(hoveron="fills") %>% 
    layout(xaxis = list(range=c(-88.3, -87.5)),
           yaxis = list(range=c(41.4, 42.2)))
  
})
```

Column
------------------

### Change in percent ridership/use

```{r}
renderPlotly({
  chicago_change <- chicago_transit %>%
    select(NAME, contains(input$transit_input_chi)) %>% 
    filter(str_detect(NAME, input$location_input_chi)) %>% 
    pivot_longer(cols = contains("change")) %>% 
    mutate(
      tooltip = NAME %>% 
        str_replace("--", "\n - ") %>% 
        str_replace(" PUMA(;|,) Illinois", "") %>% 
        str_replace_all("Townships? ?", "") %>% 
        str_replace_all(", |(?<!(North|South|East|West) )&", "\n - ")
    )
  
  p_perc <- chicago_change %>% 
    filter(str_detect(name, "perc")) %>% 
    ggplot(aes(fill = value, text = tooltip)) +
      geom_sf(color = "black", lwd = 0.1) +
      coord_sf(xlim = c(-88.3, -87.5), ylim = c(41.4, 42.2)) + 
      scale_fill_viridis_c(name = "% Use", option = "magma") +
      theme_void() +
      labs(
        title = "Change in % use"
      )
  
  ggplotly(p_perc, tooltip = "text") %>% 
    style(hoveron="fills") %>% 
    layout(xaxis = list(range=c(-88.3, -87.5)),
           yaxis = list(range=c(41.4, 42.2)))
  
})
```


Los Angeles {data-navmenu="Change in Transit Use"}
==================

Column {.sidebar}
------------------


```{r}
selectInput(
  inputId = "location_input_la",
  label = "Choose a location:",
  choices = c("Los Angeles County" = "Los Angeles County",
              "Cities of Los Angeles & Long Beach" = "(LA City|Long Beach)",
              "Suburban/Rural LA County" = "^((?!LA City|Long Beach).)*$")
)

selectInput(
  inputId = "transit_input_la",
  label = "Choose a transit type:",
  choices = c("Car" = "change_car",
              "Work from home" = "change_wfh",
              "Public transportation" = "change_transit")
)
```


Column
------------------

### Change in raw ridership/use

```{r}
renderPlotly({
  la_change <- la_transit %>%
    select(NAME, contains(input$transit_input_la)) %>% 
    filter(str_detect(NAME, input$location_input_la)) %>% 
    pivot_longer(cols = contains("change")) %>% 
    mutate(
      tooltip = NAME %>% 
        str_replace("--", "\n - ") %>% 
        str_replace(" PUMA(;|,) California", "") %>% 
        str_replace_all("Cities ?", "") %>% 
        str_replace_all(", |(?<!(North|South|East|West) )&", "\n - ")
    )
  
  p_perc <- la_change %>% 
    filter(str_detect(name, "raw")) %>% 
    ggplot(aes(fill = value, text = tooltip)) +
      geom_sf(color = "black", lwd = 0.1) +
      coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
      scale_fill_viridis_c(name = "Use", option = "magma") +
      theme_void() +
      labs(
        title = "Change in Raw Use"
      )
  
  ggplotly(p_perc, tooltip = "text") %>% 
    style(hoveron="fills") %>% 
    layout(xaxis = list(range=c(-119.0, -117.6)),
           yaxis = list(range=c(33.6, 34.9)))
  
})
```

Column
------------------

### Change in percent ridership/use

```{r}
renderPlotly({
  la_change <- la_transit %>%
    select(NAME, contains(input$transit_input_la)) %>% 
    filter(str_detect(NAME, input$location_input_la)) %>% 
    pivot_longer(cols = contains("change")) %>% 
    mutate(
      tooltip = NAME %>% 
        str_replace("--", "\n - ") %>% 
        str_replace(" PUMA(;|,) California", "") %>% 
        str_replace_all("Cities ?", "") %>% 
        str_replace_all(", |(?<!(North|South|East|West) )&", "\n - ")
    )
  
  p_perc <- la_change %>% 
    filter(str_detect(name, "perc")) %>% 
    ggplot(aes(fill = value, text = tooltip)) +
      geom_sf(color = "black", lwd = 0.1) +
      coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
      scale_fill_viridis_c(name = "% Use", option = "magma") +
      theme_void() +
      labs(
        title = "Change in % use"
      )
  
  ggplotly(p_perc, tooltip = "text") %>% 
    style(hoveron="fills") %>% 
    layout(xaxis = list(range=c(-119.0, -117.6)),
           yaxis = list(range=c(33.6, 34.9)))
  
})
```


Analysis
==================

Column
------------------

### Change in Countywide Public Transit Ridership

```{r}
county_transit %>% 
  filter(variable == "transit_users") %>% 
  ggplot(aes(x = year, y = estimate, color = NAME)) +
    geom_point() +
    geom_ribbon(aes(ymin = 0, ymax = estimate, fill = NAME), alpha = .3) +
    theme_minimal() +
    scale_x_continuous(breaks = c(2019, 2021)) +
    scale_y_continuous(limits = c(0, 500000),
                       labels = scales::label_number(suffix = "K", scale = 1e-3)) +
    scale_color_discrete(guide = "none") +
    labs(
      title = "Change in Public Transit Ridership, 2019-2021",
      subtitle = "Cook County (Chicago) sees steeper decline over pandemic\nthan Los Angeles County",
      x = "Year",
      y = "Ridership (in thousands)",
      fill = "",
      caption =  "Data from American Community Survey 2019 & 2021\nNote: Data is incomplete"
    ) +
    theme(
      legend.position = "bottom"
    )
```

Column {.tabset .tabset-fade}
------------------

### Change in Transit Ridership by PUMA

```{r}
# raw ridership
# Chicago
chi_raw <- chicago_transit %>% 
  select(NAME, contains("transit_users")) %>% 
  pivot_longer(cols = contains("transit_users")) %>% 
  mutate(name = str_replace_all(name, "transit_users_", "")) %>% 
  ggplot(aes(fill = value)) +
    geom_sf(color = "black") +
    facet_wrap(~name) +
    scale_fill_viridis_c(name = "Public transit\nridership", option = "magma") +
    theme_void() +
    labs(
      subtitle = "Cook County, IL (Chicago)"
    )

# Los Angeles
la_raw <- la_transit %>% 
  select(NAME, contains("transit_users")) %>% 
  pivot_longer(cols = contains("transit_users")) %>%
  mutate(name = str_replace_all(name, "transit_users_", "")) %>%
  ggplot(aes(fill = value)) +
    geom_sf(color = "black") +
    facet_wrap(~name) +
    scale_fill_viridis_c(name = "Public transit\nridership", option = "magma") +
    coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
    theme_void() +
    labs(
      subtitle = "Los Angeles County, CA"
    )

raw_plot <- ggarrange(chi_raw, la_raw, ncol = 1, nrow = 2)

annotate_figure(raw_plot, 
                top = text_grob("Number of Workers Using Public Transit\nto Commute in Chicago and LA, 2019-2021",
                                face = "bold",
                                size = 12))
```

### Change in Percent Transit Ridership by PUMA

```{r}
chi_perc <- chicago_transit %>% 
  ggplot(aes(fill = perc_change_transit)) +
    geom_sf(color = "black") +
    scale_fill_viridis_c(name = "Change in\n% transit use", option = "magma", limits = c(-40,0)) +
  theme_void() +
  labs(subtitle = "Cook County, IL (Chicago)")


la_perc <- la_transit %>% 
  ggplot(aes(fill = perc_change_transit)) +
    geom_sf(color = "black") +
    scale_fill_viridis_c(name = "Change in\n% transit use", option = "magma", limits = c(-40,0)) +
    coord_sf(xlim = c(-119.0, -117.6), ylim = c(33.6, 34.9)) +
    theme_void() +
    labs(subtitle = "Los Angeles County, CA")

perc_plot <- ggarrange(chi_perc, la_perc, ncol = 2, nrow = 1, common.legend = TRUE, legend = "right")

annotate_figure(perc_plot, 
                top = text_grob("Change in Percent of Workers Using Public Transit\nto Commute in Chicago and LA, 2019-2021",
                                face = "bold",
                                size = 12))
```


